Can Symmetric Cryptography Be Liberated from the Von Neumann Style?

### 2024-10-16 paper.tex
Confusion about diffusion

Following the NSA involvement in the design of the \DES{} diffusion
layer~\citep{johnson:cold-war-crypto}, cryptographers have been
advocating for ``nothing-up-my-sleeve'' designs (which does not make
magic tricks impossible, just
harder~\citep{bernstein:something-up-my-sleeves}). For example, the
diffusion layer of the \AES{} standard is based on the usual
multiplicative inverse over the finite field induced by an
innocent-looking polynomial. To support such design, the state modern
ciphers operate on is taken to represent a certain matrix
(oftentimes, a 2D or 3D array of bits): the cipher is not specified
over a flat sequence of individual bits but over a n-dimensional array
of bits.

>>>>

Following the NSA involvement in the design of the \DES{} confusion
layer~\citep{johnson:cold-war-crypto}, cryptographers have been
advocating for ``nothing-up-my-sleeve'' designs (which does not make
magic tricks impossible, just
harder~\citep{bernstein:something-up-my-sleeves}). For example, the
confusion layer of the \AES{} standard is based on the usual
multiplicative inverse over the finite field induced by an
innocent-looking polynomial. To support such design, the state modern
ciphers operate on is taken to represent a certain matrix
(oftentimes, a 2D or 3D array of bits): the cipher is not specified
over a flat sequence of individual bits but over a n-dimensional array
of bits.
